---
title: "Bootstrapped Confidence Interval for Proportions"
subtitle: "ESM 244"
author: "Roupen Khanjian"
date: "Winter 2021"
output: 
  html_document:
    theme: lumen
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse) # Easily Install and Load the 'Tidyverse', CRAN v1.3.0
library(here) # A Simpler Way to Find Your Files, CRAN v1.0.1
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data, CRAN v2.1.0
library(boot) # Bootstrap Functions (Originally by Angelo Canty for S), CRAN v1.3-26
library(broom) # Convert Statistical Objects into Tidy Tibbles, CRAN v0.7.4
library(gt) # Easily Create Presentation-Ready Display Tables, CRAN v0.2.2



```

## Create Proportion function and Find Bootstrap Samples

```{r}

prob_initial <- 22/36 # proportion of initial sample

# make a sample where 1 = experienced exclusionary conduct
ucsb_vector <- c(rep(0,14), rep(1,22)) 

# randomize sample (not sure if theis step is necessary)
ucsb_vector <- sample(ucsb_vector, 36, replace = FALSE)

# sum(ucsb_vector) # double check that there are 22 1s

# Create function that calculates proportion
proportions_function <- function(x,i) {
  sum(rbinom(36, 1, x[i]))/ 36
}

# Create 50,000 bootstrap samples
ucsb_boot_50k <- boot(ucsb_vector,
                        statistic = proportions_function,
                        R = 50000) # 50,000 bootstrap samples

# Table of Bootstrap stats
tidy(ucsb_boot_50k) %>% 
  gt() %>% 
  tab_header(
    title = "Bootstrap Statistics",
    subtitle = "From 50,000 bootstrap samples"
  ) %>% 
  fmt_number(
    columns = vars(bias),
    decimals = 7
  ) %>% 
  fmt_number(
    columns = vars(statistic),
    decimals = 3
  ) %>% 
  fmt_number(
    columns = vars(std.error),
    decimals = 5
  )%>% 
  cols_label(
    statistic = "Sample Mean",
    bias = "Bias",
    std.error = "Standard Error"
  )
  

```



## Histogram of Bootstrapped Proportions

```{r}
# convert vector into data frame
ucsb_boot_50k_df <- data.frame(bs_prop = ucsb_boot_50k$t)

ggplot(data = ucsb_boot_50k_df,
       aes(x = bs_prop)) +
  geom_histogram(color = "white",
                 fill = "mediumaquamarine", bins = 25) +
  geom_rug(color = "red",
           alpha = 0.1) +
  scale_x_continuous(breaks = seq(0.3, 0.9, 0.1)) +
  scale_y_continuous(breaks = seq(0, 6000, 1000)) +
  theme_minimal()  +
  labs(x = "Bootstrap Proportion",
      y = "Count") +
  theme(axis.text = element_text(face = "bold",
                                 size = 11),
        axis.title = element_text(face = "bold",
                                 size = 12.5),
        plot.caption = element_text(face = "bold",
                                  size = 14.5),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "gray",
                              fill = NA))


```



## Table Breakdown of Bootstrap Samples

```{r}

ucsb_boot_50k_df %>% 
  count(bs_prop) %>% 
  rename(Proportion = "bs_prop") %>%
  gt() %>% 
  tab_header(
    title = "Bootstrap samples"
  ) %>% 
  fmt_number(
    columns = vars(Proportion),
    decimals = 3
  ) %>% 
  data_color(
    columns = vars(n),
    colors = scales::col_numeric(
      palette = c(
        "lightskyblue1", "dodgerblue", "royalblue4") ,
      domain = NULL
      )
    )

```

## Confidence Interval

```{r}

# Obtain the 95% confidence interval
ucsb_ci <- boot.ci(ucsb_boot_50k, conf = 0.95, type = "perc")

# Put values into tibble
tibble_ci <- tibble(`Lower Bound` = ucsb_ci$percent[4], 
                    `Upper Bound` = ucsb_ci$percent[5])

# Combine with stats from bootstrap      
boot_all_stats <- tibble(tidy(ucsb_boot_50k), tibble_ci)

# Create table
boot_all_stats %>% 
  gt() %>% 
  tab_header(
    title = "Bootstrap Statistics and Confidence Interval",
    subtitle = "From 50,000 bootstrap samples"
  ) %>% 
  fmt_number(
    columns = vars(bias),
    decimals = 7
  ) %>% 
  fmt_number(
    columns = vars(statistic, `Lower Bound`, `Upper Bound`),
    decimals = 3
  ) %>% 
  fmt_number(
    columns = vars(std.error),
    decimals = 5
  )%>% 
  cols_label(
    statistic = "Sample Mean",
    bias = "Bias",
    std.error = "Standard Error"
  ) %>% 
  tab_footnote(
    footnote = "Based off 95% confidence interval",
    locations = cells_column_labels(vars(`Lower Bound`, `Upper Bound`))
  )

```

&nbsp;

In the above table we can see that the 95% confidence interval for the proportion of genderqueer individuals experiencing exclusionary conduct is (`r round(ucsb_ci$percent[4], 3)`, `r round(ucsb_ci$percent[5], 3)`). This means that we are 95% confident that the true population proportion of genderqueer individuals experiencing exclusionary conduct falls within this interval. Thus, if we created an infinite number of confidence intervals we would expect 95% of them to contain the true population proportion parameter. 

