---
title: 'Parameter Estimation: Wild Fish Catch'
subtitle: "ESM 244"
author: "Roupen Khanjian"
date: "Winter 2021"
output: 
  html_document:
    theme: cosmo
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse) # Easily Install and Load the 'Tidyverse', CRAN v1.3.0
library(here) # A Simpler Way to Find Your Files, CRAN v1.0.1
library(janitor) # Simple Tools for Examining and Cleaning Dirty Data, CRAN v2.1.0
library(lubridate) # Make Dealing with Dates a Little Easier, CRAN v1.7.9.2
library(gt) # Easily Create Presentation-Ready Display Tables, CRAN v0.2.2
library(broom) # Convert Statistical Objects into Tidy Tibbles, CRAN v0.7.4
library(nlstools) # Tools for Nonlinear Regression Analysis, CRAN v1.0-2
library(latex2exp) # Use LaTeX Expressions in Plots, CRAN v0.4.0
library(patchwork) # The Composer of Plots, CRAN v1.1.1

```


## Exploratory Graph of Wild Catch Over Time and Mathematical Estimates

```{r}

# Read in data and clean:
fish <- read_csv(here("data", "fish_catch.csv"), skip = 2) %>% # skip first 2 rows above header
  slice(-c(1:2, 66:69)) %>% # remove rows without data
  clean_names() %>% # clean column names
  mutate(wild_catch = as.numeric(wild_catch)) %>% # convert wild_catch to numeric
  mutate(year = as.numeric(year)) %>% # convert year to numeric 
  mutate(index = seq(0,  62, 1)) # add an index column for nls

# eda plot:
ggplot(data = fish,
       aes(x = year, y = wild_catch)) +
  geom_line(size = 1.75, color = "dodgerblue", alpha = 0.5) +
  geom_point(size = 2.5, color = "coral3") +
  scale_x_continuous(breaks = seq(1950, 2012, 10)) + 
  scale_y_continuous(breaks =  seq(20, 100, 10)) +
  theme_minimal() +
  labs(x = "Year",
      y = "Wild Catch (Millions of Tons)") +
  theme(axis.text = element_text(face = "bold",
                                 size = 11),
        axis.title = element_text(face = "bold",
                                 size = 12),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "gray",
                              fill = NA))


```

In the above graph we see a logistic growth model, where the global wild fish catch is increasing exponentially until it reaches a carrying capacity at around 1990. Mathematically this relationship looks like:

$$
\begin{equation}
P(t) = \frac{K}{1 + Ae^{-kt}}, \text{where} \\
K = \text{carrying capacity} \\
k = \text{population growth rate constant} \\
A = \frac{K - P_0}{P_0}, \text{where } P_0 = \text{ inital population size} 
\end{equation}
$$

```{r}
ggplot(data = fish,
       aes(x = year, y = wild_catch)) +
  geom_point(size = 2, color = "coral3") +
  geom_hline(yintercept = 90,  linetype = "dashed", # add K line
             color = "darkslateblue", size = 1.5, alpha = 0.9) +
  annotate("text", x = 1965, y = 85, # add K text
           label = "K~90", size = 10, color = "darkslateblue") +
  geom_hline(yintercept = 20,linetype = "dashed", # add P_0 line
             color = "forestgreen", size = 1.5, alpha = 0.9) +
  annotate("text", x = 2000, y = 26, # add P_0 text
           label = unname(TeX("$P_0 ~ 20")), size = 10, color = "forestgreen") +
  scale_x_continuous(breaks = seq(1950, 2012, 10)) +
  scale_y_continuous(breaks =  seq(20, 100, 10)) +
  theme_minimal() +
  labs(x = "Year",
      y = "Wild Catch (Millions of Tons)") +
  theme(axis.text = element_text(face = "bold",
                                 size = 11),
        axis.title = element_text(face = "bold",
                                 size = 12),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "gray",
                              fill = NA))
```


Looking at the above plot will be using initial estimates of parameters as:

* $K$ ~ 90 (Millions of tons)
* $P_0$ ~ 20 (Millions of tons)
* $A$ ~ $\frac{90 - 20}{20} =$  `r (90 - 20)/20` 

To find an initial estimate for $k$, first we'll graph the log of `wild catch` and see where is our exponential growth phase

```{r}
ggplot(data = fish,
       aes(x = year, y = log(wild_catch))) +
  geom_point(size = 2, color = "coral3") +
  scale_x_continuous(breaks = seq(1950, 2012, 10)) +
  theme_minimal() +
  labs(x = "Year",
      y = "ln(Wild Catch)",
      title = "Log-Transformed Wild Catch") +
  theme(axis.text = element_text(face = "bold",
                                 size = 11),
        axis.title = element_text(face = "bold",
                                 size = 12.5),
        plot.title = element_text(face = "bold",
                                  size = 14.5),
        panel.grid.minor.x = element_blank(),
        panel.background = element_rect(color = "gray",
                              fill = NA))
```

We can define our exponential growth phase to be from 1950 until 1970. 

```{r}

# Going to define exponential growth phase as pre-1970 exp growth phase first
fish_exp <- fish %>% 
  filter(year < 1970) %>% # filter for only exponential growth phase
  mutate(ln_catch = log(wild_catch)) # add response log response column to find k

lm_k <- lm(ln_catch ~ year,
           data = fish_exp) # linear model to find slope coef

```

* $k$ ~ `r round(coef(lm_k)[2], 3)`  (Millions of tons growth per year)

## Nonlinear Least Squares to Find Parameters


```{r}

# nls model
fish_nls <- nls(wild_catch ~ K/(1 + A*exp(-k*index)), # formula
                data = fish, # data
                start = list(K = 90, A = 3.5, k = 0.058), # initial estimates of parameters
                trace = FALSE) # no trace output of iterations

model_fish <- tidy(fish_nls)

# nice looking gt table
model_fish[1:2] %>% 
  gt() %>% 
  tab_header( # add header
    title = "Parameter estimates",
    subtitle = "Logistic Growth Model"
  ) %>% 
  fmt_number( # convert to numeric column
    columns = vars(estimate),
    decimals = 4
  ) %>% 
  tab_style(
    style = list(
      cell_text(style = "italic"),
      cell_borders(
        side = c("right"), 
        color = "black",
        weight = px(2)
        )
    ),
    locations = cells_body(
      columns = 1
    )
  ) %>% 
  cols_label(
    estimate = "Estimate (Millions of tons)"
  )
  

```


## Finalized graph with Data and Model Output

```{r}

# Find predicted value from model:
p_predict <- predict(fish_nls)

# Bind predictions to original data frame:
fish_complete <- data.frame(fish, p_predict)

# Plot them all together:
ggplot(data = fish_complete, aes(x = year, y = wild_catch)) +
  geom_point(size = 2, color = "coral3") +
  geom_line(aes(x = year, y = p_predict), color = "dodgerblue1",
            size = 2, alpha = 0.7) +
    annotate("text", x = 2002, y = 70.3,
             label = unname(TeX("$P(t) = \\frac{100.28}{1 + 4.32e^{-0.07t}}$")), 
             size = 7, color = "dodgerblue1") +
  scale_x_continuous(breaks = seq(1950, 2012, 10)) +
  scale_y_continuous(breaks =  seq(20, 100, 10)) +
  theme_minimal() +
  labs(x = "Year",
      y = "Wild Catch (Millions of Tons)") +
  theme(axis.text = element_text(face = "bold",
                                 size = 11),
        axis.title = element_text(face = "bold",
                                 size = 12.5),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(color = "gray",
                              fill = NA))

```



